---
title: "SurvivalCox"
author: "Nionella Bt Stephen Sampil"
date: "2023-12-21"
output: html_document
---

# Introduction

## Group Members

Nionella binti Stephen Sampil

Wan Nor Syafiqah binti Wan Salleh

Nur Hafizah binti Sukeri

Farah Munirah binti Mior Mazlan

Zahid bin Zulkifli

Ahmad Firdaus bin Mohamed

## Dataset

The dataset represent data from the Framingham Heart Study, Levy(1999) National Heart Lung and Blood Institute, Center for Bio-medical Communication.The dataset consist of measurements of 9 variables on 4699 patients who were free of coronary heart disease at their baseline exam.

The variables are:

1.  id - patient identifier
2.  Sex - Patient gender (categorical; 1 = male, 2 = female)
3.  SBP - Systolic blood pressure, mmHg (numerical)
4.  DBP - Diastolic Blood Pressure, mmHg (numerical)
5.  SCL - Serum Cholesterol, mg/dL (numerical)
6.  Age - Age at baseline exam (years) (numerical)
7.  BMI - Body Mass Index (kg/m2) (numerical)
8.  Month - Month of year at baseline exam (numerical)
9.  Follow-up - Subject's follow-up, days since baseline (numerical)
10. CHDfate - Event of CHD at end of follow-up (categorical ; 1 = chd, 0 = no chd)

# Method

The objective of the Framingham Heart Study was to identify the common factors or characteristics that contribute to risk of developing CVD by following its development over a long period of time in a large group of participants who had not yet developed overt symptoms of CVD or suffered a heart attack or stroke by cox proportional hazard regression. We can also analyze the time to event analysis (event of interest is event of developed CHD at end of follow-up)

### Prepare environment / Load libraries

```{r}
library(haven)
library(tidyverse)
library(gtsummary)
library(broom)
library(knitr)
library(tidyr)
library(survival)
library(corrplot)
```

### Read Data

```{r}
Data1 <- read_sav("FraminghamLabel.sav")
summary(Data1)
glimpse(Data1)
```

### Transform data

```{r}
Data2 <-
  Data1 %>% mutate(across(where(is.labelled), as_factor)) %>%
    mutate(scl = as.integer(scl), bmi = as.integer(bmi))
summary(Data2)
glimpse(Data2)
```

```{r}
daystime <- Data2$followup
daystime
```

```{r}
# Omit rows with missing values
Data3 <- na.omit(Data2)
```

### Describe data

Describe data for numerical variables, and categorical varibles.

Event: CHD fate time: daystime

CV:

Numerical variables: SBP, DBP, SCL, Age, BMI Categorical variables:

```{r}
Data3 %>% group_by(chdfate) %>%
    summarise(mean.sbp = mean(sbp), sd.age = sd(sbp), 
              mean.dbp = mean(dbp), sd.sbp = sd(dbp),
              mean.scl = mean(scl), sd.scl = sd(scl),
              mean.age = mean(age), sd.age = sd(age),
              mean.bmi = mean(bmi), sd.bmi = sd(bmi))
```

### All variables (IMPORTANT TO REMEMEBER!!!)

```{r}
Data3 %>% 
  tbl_summary(by = chdfate, statistic = list(all_continuous() ~ "{mean}, ({sd})", all_categorical() ~ "{n} /{N} ({p}%)"))
```

```{r}
str(Data3)
```

## Kaplan-Meir Survival Estimates

### KM Estimates for overall

Estimate the survival probabilities for all subjects (constant \~ 1)

```{r}
KM1 <- survfit(Surv(time = followup, event = chdfate == 'CHD') ~ 1, type = "kaplan-meier", data = Data3)
summary(KM1)
```

Manual calculations of survival probabilities:

371 4621 1 0.992 0.001283 0.990 0.995 374 4620 1 0.992 0.001301 0.990 0.995

At 374 days, the survival probabilities is:

```{r}
1-1/4620
```

```{r}
0.9997835*0.992
```

#### Survival Plot for overall

```{r}
ggsurvplot(KM1, data = Data3, risk.table = TRUE, linetype = c(1,2), pval = TRUE)
```

#### KM Estimates for groups

Estimate the survival probabilities by groups (categorical variables)

By Gender only in this dataset (Male and Female)

```{r}
KM.sex <- survfit(Surv(time = followup, event = chdfate == 'CHD') ~ sex, type = "kaplan-meier", data = Data3)
summary(KM.sex)
```

```{r}
ggsurvplot(KM.sex, data = Data3, linetype = c(1,2), pval = TRUE)
```

There is difference in survival between male and female groups

#### Estimate Survival function

Estimate the survival function at any percentile.

Example, what is the value for survival duration (days) at 25, 50 & 75 percentile?

```{r}
quantile_points <- c(0.25, 0.50, 0.75)
survival_times <- quantile(KM1$time, quantile_points)
summary(survival_times)
```

```{r}
quantile(KM1, probs = c(0.25, 0.50, 0.75))
```

```{r}
quantile(KM.sex, probs = c(0.25,0.50, 0.75))
```

#### Estimate survival probability

What is the minimum duration and max. duration of followup time according to chdfate (chd vs no chd)

```{r}
Data3 %>% group_by(chdfate) %>%
  summarize(min.dur = min(followup), max.dur = max(followup))
```

Lets estimate survival probability at 10, 1000, 10000 days

```{r}
summary(KM1, times = c(10, 1000, 10000))
```

2nd example, lets ectimate the survival probabilities at 5 years (1825days), 10 years(3650days), and 15 years(5475)

```{r}
summary(KM1, times = c(1825, 3650, 5475))
```

3rd example, by using KM.sex to estimate survival probabilities for femala at 5 years (1825days), 10 years(3650days), and 15 years(5475day. s)

```{r}
summary(KM.sex, times = c(1825, 3650, 5475))
```


### Compare survival estimates between categorical variable

We need to compare the survival estimates of gender group(female vs males) y performing logrank test, peto test, weilbull test etc.

#### logrank test

```{r}
logrank.sex <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ sex, data = Data3, rho = 0)
logrank.sex
```


#### Peto test

```{r}
peto.sex <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ sex, data = Data3, rho = 1)
peto.sex
```

From both logrank test and peto test, the P value is <0.001, hence there is statistically significant different of survival estimates between female and male group.



### Compare survival estimates between numerical groups 

Numerical variable:
Age, BMI, SCL, SBP, DBP

#### Age

```{r}
summary(Data3$age)
```
Categorize age into groups:

0, 30, 45, 70 

```{r}
Data3 <- Data3 %>% 
  mutate(age.c = cut(age, c(0, 30, 45, 70), labels = c('min-30', '31-45', 'above 45')))
Data3 %>% count(age.c)
```

```{r}
KM.age <- survfit(Surv(time = followup, chdfate == 'CHD') ~ age.c, data = Data3, type = "kaplan-meier")
summary(KM.age)
```

```{r}
ggsurvplot(KM.age, data = Data3, linetype = c(1, 2, 3), risk.table = TRUE, pval = TRUE)
```




```{r}
logrank.age.c <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ age.c, data = Data3, rho = 0)
logrank.age.c
```
```{r}
peto.age.c <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ age.c, data = Data3, rho = 1)
peto.age.c
```
#### BMI

```{r}
summary(Data3$bmi)
```
Categorize into:

0, 15, 25, 40, 60

```{r}
Data3 <- Data3 %>% mutate(bmi.c = cut(bmi, c(0, 15, 25, 40, 60), labels = c('min-15', '16-25', '26-40','above 40')))
Data3 %>% count(bmi.c)
```

```{r}
KM.bmi <- survfit(Surv(time = followup, event = chdfate == 'CHD') ~ bmi.c, type = "kaplan-meier", data = Data3)
summary(KM.bmi)
```

```{r}
ggsurvplot(KM.bmi, data = Data3, risk.table = TRUE, linetype = c(1,2,3), pval = TRUE)
```



```{r}
logrank.bmi <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ bmi.c, data = Data3, rho = 0)
logrank.bmi
```
```{r}
peto.bmi <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ bmi.c, data = Data3, rho = 1)
peto.bmi
```




#### SBP

```{r}
summary(Data3$sbp)
```

Categorize into:

0, 110, 130, 300

```{r}
Data3 <- Data3 %>% mutate(sbp.c = cut(sbp, c(0, 110, 130, 300), labels = c('min-110', '111-130','above 130')))
Data3 %>% count(sbp.c)
```

```{r}
KM.sbp <- survfit(Surv(time = followup, event = chdfate == 'CHD') ~ sbp.c, type = "kaplan-meier", data = Data3)
summary(KM.sbp)
```


```{r}
ggsurvplot(KM.sbp, data = Data3, risk.table = TRUE, linetype = c(1,2,3), pval = TRUE)
```





```{r}
logrank.sbp <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ sbp.c, data = Data3, rho = 0)
logrank.sbp
```


```{r}
peto.sbp <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ sbp.c, data = Data3, rho = 1)
peto.sbp
```




#### DBP

```{r}
summary(Data3$dbp)
```
categorize into:

0, 70, 90, 150

```{r}
Data3<- Data3 %>%
  mutate(dbp.c = cut(dbp, c(0, 70, 90, 150), labels = c('min-70', '71-90', ' above 90')))
Data3 %>% count(dbp.c)
```

```{r}
KM.dbp <- survfit(Surv(time = followup, event = chdfate == 'CHD') ~ dbp.c, type = "kaplan-meier", data = Data3)
summary(KM.dbp)
```

```{r}
ggsurvplot(KM.dbp, data = Data3, risk.table = TRUE, linetype = c(1,2,3), pval = TRUE)
```





```{r}
logrank.dbp <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ dbp.c, data = Data3, rho = 0)
logrank.dbp
```

```{r}
peto.dbp <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ dbp.c, data = Data3, rho = 1)
peto.dbp
```





#### SCL

```{r}
summary(Data3$scl)
```

Categorize into:

0, 200, 255, 600


```{r}
Data3<- Data3 %>%
  mutate(scl.c = cut(scl, c(0, 200, 255, 600), labels = c('min-200', '201-255', ' above 255')))
Data3 %>% count(scl.c)
```

```{r}
KM.scl <- survfit(Surv(time = followup, event = chdfate == 'CHD') ~ scl.c, type = "kaplan-meier", data = Data3)
summary(KM.scl)
```

```{r}
ggsurvplot(KM.scl, data = Data3, risk.table = TRUE, linetype = c(1,2,3), pval = TRUE)
```




```{r}
logrank.scl <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ dbp.c, data = Data3, rho = 0)
logrank.scl
```

```{r}
peto.scl <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ scl.c, data = Data3, rho = 1)
logrank.scl
```


# COX PROPORTIONAL HAZARD REGRESSION (SIMPLE/UNIVARIATE)


Estimate the COX PH regression using Elfron method or Breslow method. 

Outcome variable:

Time : followup
Event: chdfate

Covariates:
Numerical: age, sbp, dbp, scl, bmi 
Categorical : gender 

## Univariable COX PH regression 


### Numerical variable 

#### Age

##### Elfron method

```{r}
cox.age <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ age, data = Data3)
summary(cox.age)
```





##### Breslow method 

```{r}
cox.age.breslow <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ age, data = Data3, method = "breslow")
summary(cox.age.breslow)
```



#### SBP


```{r}
cox.sbp <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ sbp, data = Data3)
summary(cox.sbp)
```

```{r}
cox.sbp.breslow <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ sbp, data = Data3, method = "breslow")
summary(cox.sbp.breslow)
```



#### DBP


```{r}
cox.dbp <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ dbp, data = Data3)
summary(cox.dbp)
```

```{r}
cox.dbp.breslow <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ dbp, data = Data3, method = "breslow")
summary(cox.dbp.breslow)
```



#### SCL

```{r}
cox.scl <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ scl, data = Data3)
summary(cox.scl)
```


```{r}
cox.scl.breslow <-coxph(Surv(time = followup, event = chdfate == 'CHD') ~ scl, data = Data3, method = "breslow")
summary(cox.scl.breslow)
```



#### BMI

```{r}
cox.bmi <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ bmi, data = Data3)
summary(cox.bmi)
```


### Categorical variable 

#### Gender 

```{r}
cox.sex <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ sex, data = Data3)
summary(cox.sex)
```

#### age.c, sbp.c, bmi.c

```{r}
cox.age.c <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ age.c, data = Data3)
cox.sbp.c <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ sbp.c, data = Data3)
cox.bmi.c <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ bmi.c, data = Data3)
```


```{r}
summary(cox.age.c)
summary(cox.sbp.c)
summary(cox.bmi.c)
```
 

### ALL VARIABLE (Table summary)****


```{r}
Data3 %>% 
  dplyr::select(-age.c) %>%
  tbl_uvregression(method = coxph,
  y = Surv(time = followup, event = chdfate == 'CHD'),
  exponentiate = TRUE,
  pvalue_fun = ~style_pvalue(.x, digits = 3)
  ) %>% bold_labels() %>% italicize_levels() %>%
  as_gt()
```



## Multivariable Cox PH Regression 


### Main effect model

Outcome variable:

Time : followup
Event: chdfate

Covariates:
Numerical: age, sbp, dbp, scl, bmi 
Categorical : gender 

```{r}
mv.cox <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ age + sbp + dbp + scl + bmi + sex, data = Data3)
summary(mv.cox)
tidy(mv.cox)     # log hazard
tidy(mv.cox, exponentiate = TRUE)      #hazard ratio
```



### Model with interaction

? Numerical vs Numerical (age vs bmi), as age increases, bmi may also increase //scl vs bmi
? Numerical vs Categorical (gender vs scl), as women has more cholesterol levels comapred to men // 
 
 

#### Age vs BMI 


```{r}
mv.cox.age.bmi.ia <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ age + sbp + dbp + scl + bmi + sex + age:bmi, data = Data3)
summary(mv.cox.age.bmi.ia)
tidy(mv.cox.age.bmi.ia)    # Log Hazards
tidy(mv.cox.age.bmi.ia, exponentiate = TRUE)    # Hazard ratio
```


#### Gender vs SCL


```{r}
mv.cox.sex.scl.ia <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ age + sbp + dbp + scl + bmi + sex + sex:scl, data = Data3)
summary(mv.cox.sex.scl.ia)
tidy(mv.cox.sex.scl.ia)
tidy(mv.cox.sex.scl.ia, exponentiate = TRUE)
```

#### SCL vs BMI 


```{r}
mv.cox.scl.bmi.ia <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ age + sbp + dbp + scl + bmi + sex + scl:bmi, data = Data3)
summary(mv.cox.scl.bmi.ia)
tidy(mv.cox.scl.bmi.ia)
tidy(mv.cox.scl.bmi.ia, exponentiate = TRUE)
```
No interaction noted in model with interactions


### Model comparison

Compare main effect model with 2 model with interaction previously 

```{r}
anova(mv.cox, mv.cox.age.bmi.ia, test = 'Chisq')
```
There are no interaction between age and bmi, P value > 0.05, not significant as well

```{r}
anova(mv.cox, mv.cox.scl.bmi.ia, test = 'Chisq')
```
There are no interaction between serum cholesterol and bmi, P value > 0.05, not significant as well

```{r}
anova(mv.cox, mv.cox.sex.scl.ia, test = 'Chisq')
```

There are no interaction between serum cholesterol and gender, P value > 0.05, not significant as well














