---
title: "SurvivalCox"
author: "Nionella Bt Stephen Sampil"
date: "2023-12-21"
output: html_document
---

# Introduction

## Group Members

Nionella binti Stephen Sampil

Wan Nor Syafiqah binti Wan Salleh

Nur Hafizah binti Sukeri

Farah Munirah binti Mior Mazlan

Zahid bin Zulkifli

Ahmad Firdaus bin Mohamed

## Dataset

The dataset represent data from the Framingham Heart Study, Levy(1999) National Heart Lung and Blood Institute, Center for Bio-medical Communication.The dataset consist of measurements of 9 variables on 4699 patients who were free of coronary heart disease at their baseline exam. This dataset can be downloadable from the website https://www.causeweb.org/tshs/framingham-didactic/.

The variables are:

1.  id - patient identifier
2.  Sex - Patient gender (categorical; 1 = male, 2 = female)
3.  SBP - Systolic blood pressure, mmHg (numerical)
4.  DBP - Diastolic Blood Pressure, mmHg (numerical)
5.  SCL - Serum Cholesterol, mg/dL (numerical)
6.  Age - Age at baseline exam (years) (numerical)
7.  BMI - Body Mass Index (kg/m2) (numerical)
8.  Month - Month of year at baseline exam (numerical)
9.  Follow-up - Subject's follow-up, days since baseline (numerical)
10. CHDfate - Event of CHD at end of follow-up (categorical ; 1 = chd, 0 = no chd)

# Method

The objective is to analyze the time to event analysis (event of interest is event of developed CHD at end of follow-up).

### Prepare environment / Load libraries

```{r}
library(haven)
library(tidyverse)
library(gtsummary)
library(broom)
library(knitr)
library(tidyr)
library(survival)
library(corrplot)
library(survminer)
```

### Read Data

```{r}
Data1 <- read_sav("FraminghamCox.sav")
summary(Data1)
glimpse(Data1)
```

### Transform data

```{r}
Data2 <-
  Data1 %>% mutate(across(where(is.labelled), as_factor)) %>%
  mutate(bmi = as.numeric(bmi)) 
summary(Data2)
glimpse(Data2)
str(Data2)
```


### Describe data

Describe data for numerical variables, and categorical varibles.

Event: CHD fate
Time: followup

CV:

Numerical variables: SBP, DBP, SCL, Age, BMI Categorical variables:

```{r}
Data2 %>% group_by(chdfate) %>%
    summarise(mean.sbp = mean(sbp), sd.age = sd(sbp), 
              mean.dbp = mean(dbp), sd.sbp = sd(dbp),
              mean.scl = mean(scl), sd.scl = sd(scl),
              mean.age = mean(age), sd.age = sd(age),
              mean.bmi = mean(bmi), sd.bmi = sd(bmi))
```

### All variables (IMPORTANT TO REMEMEBER!!!)

```{r}
Data2 %>% 
  tbl_summary(by = chdfate, statistic = list(all_continuous() ~ "{mean}, ({sd})", all_categorical() ~ "{n} /{N} ({p}%)"))
```

```{r}
str(Data2)
```

### Checking multicollinearity

```{r}
Data3 <- 
  Data2 %>% 
  select(where(is.numeric))
```


```{r}
cor.Data3 <-
  cor(Data3, use = "complete.obs", method = "pearson")
head(round(cor.Data3,2))
```

```{r}
corrplot(cor.Data3, type = 'upper', order = 'hclust')
```

sbp and dbp indicates a high collinearity. It would be suggested to remove one of them in the COX PH regression model




## Kaplan-Meir Survival Estimates

### KM Estimates for overall

Estimate the survival probabilities for all subjects (constant \~ 1)

```{r}
KM1 <- survfit(Surv(time = followup, event = chdfate == 'CHD') ~ 1, type = "kaplan-meier", data = Data2)
summary(KM1)
```

Manual calculations of survival probabilities:


371 4621 1 0.992 0.001283 0.990 0.995 
374 4620 1 0.992 0.001301 0.990 0.995

At 374 days, the survival probabilities is:

```{r}
1-1/4620
```

```{r}
0.9997835*0.992
```

#### Survival Plot for overall

```{r}
ggsurvplot(KM1, data = Data2, risk.table = TRUE, linetype = c(1,2), pval = TRUE)
```

#### KM Estimates for groups

Estimate the survival probabilities by groups (categorical variables)

By Gender only in this dataset (Male and Female)

```{r}
KM.sex <- survfit(Surv(time = followup, event = chdfate == 'CHD') ~ sex, type = "kaplan-meier", data = Data2)
summary(KM.sex)
```

```{r}
ggsurvplot(KM.sex, data = Data2, linetype = c(1,2), pval = TRUE)
```

There is difference in survival between male and female groups

#### Estimate Survival function

Estimate the survival function at any percentile.

Example, what is the value for survival duration (days) at 25, 50 & 75 percentile?

```{r}
quantile_points <- c(0.25, 0.50, 0.75)
survival_times <- quantile(KM1$time, quantile_points)
summary(survival_times)
```

```{r}
quantile(KM1, probs = c(0.25, 0.50, 0.75))
```

```{r}
quantile(KM.sex, probs = c(0.25,0.50, 0.75))
```

#### Estimate survival probability

What is the minimum duration and max. duration of followup time according to chdfate (chd vs no chd)

```{r}
Data2 %>% group_by(chdfate) %>%
  summarize(min.dur = min(followup), max.dur = max(followup))
```

Lets estimate survival probability at 10, 1000, 10000 days

```{r}
summary(KM1, times = c(10, 1000, 10000))
```

2nd example, lets ectimate the survival probabilities at 5 years (1825days), 10 years(3650days), and 15 years(5475)

```{r}
summary(KM1, times = c(1825, 3650, 5475))
```

3rd example, by using KM.sex to estimate survival probabilities for femala at 5 years (1825days), 10 years(3650days), and 15 years(5475day. s)

```{r}
summary(KM.sex, times = c(1825, 3650, 5475))
```


### Compare survival estimates between categorical variable

We need to compare the survival estimates of gender group(female vs males) y performing logrank test, peto test, weilbull test etc.

#### logrank test

```{r}
logrank.sex <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ sex, data = Data2, rho = 0)
logrank.sex
```


#### Peto test

```{r}
peto.sex <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ sex, data = Data2, rho = 1)
peto.sex
```

From both logrank test and peto test, the P value is <0.001, hence there is statistically significant different of survival estimates between female and male group.



### Compare survival estimates between numerical groups 

Numerical variable:
Age, BMI, SCL, SBP, DBP

#### Age

```{r}
summary(Data2$age)
```
Categorize age into groups:

0, 30, 45, 70 

```{r}
Data4 <- Data2 %>% 
  mutate(age.c = cut(age, c(0, 30, 45, 70), labels = c('min-30', '31-45', 'above 45')))
Data4 %>% count(age.c)
```

```{r}
KM.age <- survfit(Surv(time = followup, chdfate == 'CHD') ~ age.c, data = Data4, type = "kaplan-meier")
summary(KM.age)
```

```{r}
ggsurvplot(KM.age, data = Data4, linetype = c(1, 2, 3), risk.table = TRUE, pval = TRUE)
```




```{r}
logrank.age.c <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ age.c, data = Data4, rho = 0)
logrank.age.c
```
```{r}
peto.age.c <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ age.c, data = Data4, rho = 1)
peto.age.c
```
#### BMI

```{r}
summary(Data2$bmi)
```
Categorize into:

0, 15, 25, 40, 60

```{r}
Data4 <- Data2 %>% mutate(bmi.c = cut(bmi, c(0, 15, 25, 40, 60), labels = c('min-15', '16-25', '26-40','above 40')))
Data4 %>% count(bmi.c)
```

```{r}
KM.bmi <- survfit(Surv(time = followup, event = chdfate == 'CHD') ~ bmi.c, type = "kaplan-meier", data = Data4)
summary(KM.bmi)
```

```{r}
ggsurvplot(KM.bmi, data = Data4, risk.table = TRUE, linetype = c(1,2,3), pval = TRUE)
```



```{r}
logrank.bmi <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ bmi.c, data = Data4, rho = 0)
logrank.bmi
```
```{r}
peto.bmi <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ bmi.c, data = Data4, rho = 1)
peto.bmi
```




#### SBP

```{r}
summary(Data2$sbp)
```

Categorize into:

0, 110, 130, 300

```{r}
Data4 <- Data2 %>% mutate(sbp.c = cut(sbp, c(0, 110, 130, 300), labels = c('min-110', '111-130','above 130')))
Data4 %>% count(sbp.c)
```

```{r}
KM.sbp <- survfit(Surv(time = followup, event = chdfate == 'CHD') ~ sbp.c, type = "kaplan-meier", data = Data4)
summary(KM.sbp)
```


```{r}
ggsurvplot(KM.sbp, data = Data4, risk.table = TRUE, linetype = c(1,2,3), pval = TRUE)
```





```{r}
logrank.sbp <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ sbp.c, data = Data4, rho = 0)
logrank.sbp
```


```{r}
peto.sbp <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ sbp.c, data = Data4, rho = 1)
peto.sbp
```




#### DBP

```{r}
summary(Data2$dbp)
```
categorize into:

0, 70, 90, 150

```{r}
Data4 <- Data2 %>%
  mutate(dbp.c = cut(dbp, c(0, 70, 90, 150), labels = c('min-70', '71-90', ' above 90')))
Data4 %>% count(dbp.c)
```

```{r}
KM.dbp <- survfit(Surv(time = followup, event = chdfate == 'CHD') ~ dbp.c, type = "kaplan-meier", data = Data4)
summary(KM.dbp)
```

```{r}
ggsurvplot(KM.dbp, data = Data4, risk.table = TRUE, linetype = c(1,2,3), pval = TRUE)
```





```{r}
logrank.dbp <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ dbp.c, data = Data4, rho = 0)
logrank.dbp
```

```{r}
peto.dbp <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ dbp.c, data = Data4, rho = 1)
peto.dbp
```





#### SCL

```{r}
summary(Data2$scl)
```

Categorize into:

0, 200, 255, 600


```{r}
Data4 <- Data2 %>%
  mutate(scl.c = cut(scl, c(0, 200, 255, 600), labels = c('min-200', '201-255', ' above 255')))
Data4 %>% count(scl.c)
```

```{r}
KM.scl <- survfit(Surv(time = followup, event = chdfate == 'CHD') ~ scl.c, type = "kaplan-meier", data = Data4)
summary(KM.scl)
```

```{r}
ggsurvplot(KM.scl, data = Data4, risk.table = TRUE, linetype = c(1,2,3), pval = TRUE)
```




```{r}
logrank.scl <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ dbp.c, data = Data4, rho = 0)
logrank.scl
```

```{r}
peto.scl <- survdiff(Surv(time = followup, event = chdfate == 'CHD') ~ scl.c, data = Data4, rho = 1)
logrank.scl
```


# COX PROPORTIONAL HAZARD REGRESSION (SIMPLE/UNIVARIATE)


Estimate the COX PH regression using Elfron method or Breslow method. 

Outcome variable:

Time : followup
Event: chdfate

Covariates:
Numerical: age, sbp, dbp, scl, bmi 
Categorical : gender 

## Univariable COX PH regression 


### Numerical variable 

#### Age

##### Elfron method

```{r}
cox.age <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ age, data = Data2)
summary(cox.age)
```





##### Breslow method 

```{r}
cox.age.breslow <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ age, data = Data2, method = "breslow")
summary(cox.age.breslow)
```



#### SBP


```{r}
cox.sbp <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ sbp, data = Data2)
summary(cox.sbp)
```

```{r}
cox.sbp.breslow <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ sbp, data = Data2, method = "breslow")
summary(cox.sbp.breslow)
```



#### DBP


```{r}
cox.dbp <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ dbp, data = Data2)
summary(cox.dbp)
```

```{r}
cox.dbp.breslow <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ dbp, data = Data2, method = "breslow")
summary(cox.dbp.breslow)
```



#### SCL

```{r}
cox.scl <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ scl, data = Data2)
summary(cox.scl)
```


```{r}
cox.scl.breslow <-coxph(Surv(time = followup, event = chdfate == 'CHD') ~ scl, data = Data2, method = "breslow")
summary(cox.scl.breslow)
```



#### BMI

```{r}
cox.bmi <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ bmi, data = Data2)
summary(cox.bmi)
```


### Categorical variable 

#### Gender 

```{r}
cox.sex <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ sex, data = Data2)
summary(cox.sex)
```

#### age.c, sbp.c, bmi.c

```{r}
cox.age.c <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ age.c, data = Data4)
cox.sbp.c <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ sbp.c, data = Data4)
cox.bmi.c <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ bmi.c, data = Data4)
```


```{r}
summary(cox.age.c)
summary(cox.sbp.c)
summary(cox.bmi.c)
```
 

### ALL VARIABLE (Table summary)****


```{r}
Data2 %>% 
  dplyr::select(-age.c) %>%
  tbl_uvregression(method = coxph,
  y = Surv(time = followup, event = chdfate == 'CHD'),
  exponentiate = TRUE,
  pvalue_fun = ~style_pvalue(.x, digits = 3)
  ) %>% bold_labels() %>% italicize_levels() %>%
  as_gt()
```



## Multivariable Cox PH Regression 


### Main effect model

Outcome variable:

Time : followup
Event: chdfate

Covariates:
Numerical: age, sbp, scl, bmi 
Categorical : gender 

```{r}
mv.cox <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ age + sbp + scl + bmi + sex, data = Data2)
summary(mv.cox)
tidy(mv.cox)     # log hazard
tidy(mv.cox, exponentiate = TRUE)      #hazard ratio
```



### Model with interaction

? Numerical vs Numerical (age vs bmi), as age increases, bmi may also increase //scl vs bmi
? Numerical vs Categorical (gender vs scl), as women has more cholesterol levels comapred to men // 
 
 

#### Age vs BMI 


```{r}
mv.cox.age.bmi.ia <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ age + sbp + scl + bmi + sex + age:bmi, data = Data2)
summary(mv.cox.age.bmi.ia)
tidy(mv.cox.age.bmi.ia)    # Log Hazards
tidy(mv.cox.age.bmi.ia, exponentiate = TRUE)    # Hazard ratio
```


#### Gender vs SCL


```{r}
mv.cox.sex.scl.ia <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ age + sbp + scl + bmi + sex + sex:scl, data = Data2)
summary(mv.cox.sex.scl.ia)
tidy(mv.cox.sex.scl.ia)
tidy(mv.cox.sex.scl.ia, exponentiate = TRUE)
```

#### SCL vs BMI 


```{r}
mv.cox.scl.bmi.ia <- coxph(Surv(time = followup, event = chdfate == 'CHD') ~ age + sbp + scl + bmi + sex + scl:bmi, data = Data2)
summary(mv.cox.scl.bmi.ia)
tidy(mv.cox.scl.bmi.ia)
tidy(mv.cox.scl.bmi.ia, exponentiate = TRUE)
```
There is no interaction noted in COX PH model (all model with interaction), p-value > 0.05


### Model comparison

Compare main effect model with 2 model with interaction previously 

```{r}
anova(mv.cox, mv.cox.age.bmi.ia, test = 'Chisq')
```
There are no interaction between age and bmi, P value = 0.05, not significant as well

```{r}
anova(mv.cox, mv.cox.scl.bmi.ia, test = 'Chisq')
```
There are no interaction between serum cholesterol and bmi, P value > 0.05, not significant as well

```{r}
anova(mv.cox, mv.cox.sex.scl.ia, test = 'Chisq')
```

There are no interaction between serum cholesterol and gender, P value > 0.05, not significant as well

Hence our preliminary final model is mv.cox

```{r}
prelim.final <- mv.cox
```


### Assumptions 

#### Linearity of hazard assumption

- Use mfp package to check linearity of assumption, applies to only numerical covariates



```{r}
library(mfp)
mfp.mv.cox <- mfp(Surv(time = followup, event = chdfate == 'CHD') ~ fp(age, df = 4) + fp(sbp) + fp(scl) + fp(bmi) + sex, family = cox, data = Data2, verbose = TRUE)
```
Deviance table:

Null model: The deviance associated with the null model (model with no predictors) is 23517.7. This serves as baseline for model comparison.

Linear model: The deviance associated with linear model(without any transformation) is 22747.16, a decrease in deviance indicates an improvement in model fit.

Final model: The deviance associated with final model(after applying fractional polynomials) is 22747.16. There is no further decrease from the linear model, it suggests that the specified fractional polynomial transformation did not significantly improve the model fit. 

Conclusion, the COX PH model considered **good model fit**??? to clarify



#### Proportional hazard assumption


### Constant Regression Coefficient using KM method 

```{r}
prop.h <- cox.zph(mv.cox, transform = 'km', global = TRUE)
prop.h
```
All CV p value is < 0.05 except for bmi, hence the risks are proportional overtime. Global p value is <0.05.

This result suggests that age, sbp, scl, & sex are statistically significant predictors of outcome. While bmi is not statistically significant in this model. The global p value of <0.05 indicates that the model as a whole is statistically significant.

### Constant Regression Coefficient using Rank method

```{r}
prop.h.rank <- cox.zph(mv.cox, transform = 'rank')
prop.h.rank
```


### Plots for PH Assumption


```{r}
plot(prop.h)
```



```{r}
plot(prop.h.rank)
```



## Model Checking 

Residuals:

1. Schoenfeld residuals
2. Score residuals
3. Martingale residuals
4. dfbeta residuals

### Schoenfeld residuals 


```{r}
schoen.cox <-resid(mv.cox, type = "schoenfeld")
head(schoen.cox)
```

### Score Residuals

```{r}
score.cox <- resid(mv.cox, type = "score")
head(score.cox)
```


### Martingale residuals 

```{r}
marti.cox <- resid(mv.cox, type = "martingale")
head(marti.cox)
```
#### dfbeta residuals 


```{r}
dfbeta.cox <- resid(mv.cox, type = "dfbeta")
head(dfbeta.cox)
```


### Residual plots 


To identify any outliers through residual plots 


#### Age 

```{r}
plot(Data2$age, score.cox[,1], ylab="score residuals")
```


```{r}
plot(Data2$age, marti.cox, ylab = "Martingale for Age")
```


```{r}
plot(marti.cox, type = 'h', main = "Martingale residuals", ylab = "dfbetas", lwd = 2)
```

```{r}
plot(dfbeta.cox, type = 'h', main = "dfbetas overall", ylab = "dfbetas", lwd = 2)
```

```{r}
plot(Data2$age, dfbeta.cox [,2], type = 'h', main = "Dfbetas for age", ylab = " dfbetas", lwd = 2)
```


```{r}
augment(mv.cox, data = Data2)
```
According to the residuals plots (martingale and dfbeta plots) ---> there minimal outliers noted. We concluded that the outliers do not have significant influential effect to our model, hence the outliers were not removed. 

### Influential plots

```{r}
ggcoxdiagnostics(mv.cox, type = 'deviance', linear.predictions = FALSE)
```
```{r}
ggcoxdiagnostics(mv.cox, type = 'martingale', linear.predictions = FALSE)
```
```{r}
ggcoxdiagnostics(mv.cox.age.bmi.ia, type = 'deviance', linear.predictions = FALSE)
```

```{r}
ggcoxdiagnostics(mv.cox.age.bmi.ia, type = 'martingale', linear.predictions = FALSE)
```
According to the plots above ---> there minimal outliers noted. We concluded that the outliers do not have significant influential effect to our model, hence the outliers were not removed. 


## Prediction 


We have chosen our preliminary final model (prelim.final) as our model

```{r}
summary(prelim.final)
tidy(prelim.final)     #log hazard
tidy(prelim.final, exponentiate = TRUE)   #hazard ratio
```

```{r}
summary(Data2)
```

### Create a new data frame for prediction

create a new data frame --> for observations 

Example Questions: Can we predict survival estimates for several groups?
age (40, 55, 70)
sbp (90, 130, 300)
scl (120, 220, 600)
bmi (18, 30, 60)
Sex (no, yes)


```{r}
new_data <- expand.grid(age = c(40,55,70),
                       sbp = c(90, 130, 300), 
                       scl = c(120, 220, 600),
                       bmi = c(18, 30, 60),
                       sex = c('Male', 'Female')) #### need to be VERY CAREFUL, need to follow exactly to the dataset, if not, will not produce predictions using this new data frame 
new_data
```







#### Linear Predictor for the newdata created

##### Produce Means for all predictors/ variables 

In order to manually calculate linear predictor for respective observations, need to calculate mean for:

sex                       total observation:4658
age
sbp
scl
bmi

###### mean for sex

```{r}
summary(Data2$sex)
```

```{r}
2618/4658
```

###### mean for age

```{r}
mean(Data2$age)
```

###### sbp

```{r}
mean(Data2$sbp)
```


###### scl

```{r}
mean(Data2$scl)
```


###### bmi


```{r}
mean(Data2$bmi)
```



##### Main effect model

```{r}
new_data
```


```{r}
predict.mv.cox <- predict(mv.cox, newdata = new_data, type = 'lp')
head(predict.mv.cox)
```
```{r}
augment(mv.cox, newdata = new_data)
```


Manual calculations:

How do you get the linear predictor for the first observation? 

Age: 40              mean age 46.03199
sbp: 90              mean sbp  132.7737
scl: 120             mean scl  228.2913
bmi: 18              mean bmi 25.62797
sex: 1             mean sex 0.5620438


```{r}
mv.cox$coefficients[1]*(40 - 46.03199) +
  mv.cox$coefficients[2]*(90 - 132.7737) +
  mv.cox$coefficients[3]*(120 - 228.2913) +
  mv.cox$coefficients[4]*(18 - 25.62797) +
  mv.cox$coefficients[5]*(1 - 0.5620438)
```
For 2nd observation?


** .fitted values are the estimated log hazard ratio for each individual

Example: For 1st observation who are 40 years old, has 90 Sbp, scl of 120, bmi 18 & being male , the log hazard ratio is -1.63

```{r}
exp(-1.6310371)
```

Manual calculations:

How do you get the linear predictor for the first observation? 

Age: 40              mean age 46.03199
sbp: 90              mean sbp  132.7737
scl: 120             mean scl  228.2913
bmi: 18              mean bmi 25.62797
sex: 1             mean sex 0.5620438
age:bmi             mean (age:bmi) 0.9985837 




#### Risk Score 

Calculates the relative risk (hazard ratio) of the created population against the population sample 


##### Main effect model
```{r}
risk.mv.cox <- predict(mv.cox, newdata = new_data, type = 'risk')
head(risk.mv.cox)
```
1st observation : cumulative HR of developing CHD is 0.1957265

```{r}
exp(-1.6310371)
```

The cumulative hazard risk for the 1st observation is 0.1957265


#### Expected number of event 

Estimate the expected number of event = chdfate (developing CHD) at specific point of time 

Example: 

How many individuals developed CHD at 5 years (1825days), 10 years (3650days), 15 years (5475days)?

```{r}
new_data2 <- expand.grid(chdfate = 'CHD', followup = c(1825, 3650, 5475))
new_data2
```
```{r}
new_data3 <- data.frame(new_data, new_data2)
new_data3
summary(new_data3)
```

Next we estimate the predicted number of events for these 162 individuals.

##### For main effect model 


```{r}
pred.exp <- predict(mv.cox, newdata = new_data3, type = 'expected')
pred.exp2 <- round(pred.exp, digits = 3)
cbind(new_data3, pred.exp2)

```

For the 125th observation, who are 55 years old female with initial follow up time of 10 years (3650 days) and the specific values for other covariates has predicted expected number of occurences CHD is 0.629.


27   70 300 600  18   Male     CHD     5475    24.936

for 27th observation, a 70 year old male, with initial follow up time of 15 years(5475 days) with its specific values of covariates has predicted expected number of CHD occurences is 24.936 (2493.6%).



#### Estimate survival probability using new data frame(new_data3) 

1. baseline hazard function
2. linear predictor (log hazard)

```{r}
lp.pred <- predict(mv.cox, newdata = new_data3, type = 'lp')
lp.pred     #log hazard
```


Estimate baseline hazard

```{r}
base <- basehaz(mv.cox)
base
```

140 0.0245772434 1096

What are the predicted hazard at 1096 days?

```{r}
pred.hazard <- 0.02457*exp(lp.pred)
pred.hazard
```

## Interpretation

Our final model is chosen is model with interaction (mv.cox.age.bmi.ia)

```{r}
Final.model <- prelim.final
summary(Final.model)
```
```{r}
tbl_regression(Final.model, exponentiate = TRUE, pvalue_fun = ~ style_pvalue(.x,    digits = 3)) %>%
     add_nevent(location = 'level') %>%
    bold_labels() %>% italicize_levels() %>%
    as_gt() %>% 
    gt::tab_header(title = "Table 3. Multiple Cox Proportional Hazard Regression (Final Model)")
```


#### Model equation:

Occurences of CHD (event) according to time = ho(t) * exp(1.08*age) + (1.01*sbp) + (1.01*scl) + (1.11*bmi) + (0.46*gender)


### Interpretation

1. Every 1 year increase in age, the hazard ratio of occuring CHD is 1.08.
2. Every 1 mmHg increase of systolic blood pressure, the hazard ratio of occuring CHD is 1.01 when age, serum cholesterol, BMI and gender adjusted.
3. Every 1 mg/dL increase of serum cholesterol, thee hazard ratio of CHD occuring is 1.01 when adjusted to age, systolic blood pressure, BMI and gender.

4. Every 1 unit increase of BMI, the hazard ratio of CHD occuring is 1.04 when adjusted to age, systolic blood pressure, serum cholesterol, and gender

5. Women has 55% hazard reduction of occuring CHD, compared to men when adjusted to age, systolic blood pressure, serum cholesterol, and BMI



